### 2023-07-16 supplement.qmd

This document shows the output of all our supplementary analyses.

>>>>

This document shows the output of all our supplementary analyses including data cleaning. To get the raw data run `make data`.

```{r}
#| label: packages
#| code-summary: R Packages
#| cache: false

library(patchwork)
library(gt)
library(modelr)
library(scales)
library(broom)
library(ggdist)
library(distributional)
library(tidyverse)

theme_set(
  theme_linedraw() +
    theme(
      strip.background = element_rect(fill = NA, color = NA), 
      strip.text = element_text(colour = "black", hjust = 0), 
      panel.grid = element_blank()
    )
)
```

# Data cleaning

We first load the relevant variables from the raw data file. We also rescale the raw mood reports to proportions.

```{r}
#| label: data-read
#| message: false

dat <- read_csv(
  "data/study_prompt_answered.csv",
  col_select = c(
    pid, 
    time = Time_utc, 
    duration = CurrentSessionLength,
    prompt = LastStudyPromptType, 
    mood = response
  )
) |> 
  mutate(
    pid = factor(pid),
    mood = mood / 1000
  )
dat
```

We then create a variable that identifies unique sessions, using the heuristics indicated in the code below

```{r}
#| label: clean-sessions
#| code-fold: show
#| include: true

# Create sessions
dat <- dat |> 
  arrange(pid, time) |> 
  mutate(
    i = row_number(),
    # Observation is in a new session if this...
    new_session = 
      # is the first total observation,
      i == 1 |
      # session duration is zero,
      (duration == 0) | 
      # session duration is less than previous session duration (ticker has reset),
      (duration < lag(duration, default = 0)) | 
      # or if more than 30 minutes since previous observation.
      (time > lag(time) + minutes(30)),
    # Session number is the cumulative sum of new sessions...
    session = cumsum(new_session),
    # ... per person
    .by = pid
  ) |> 
  select(-c(i, new_session))
dat
```

We filter out all other reports except mood ("Wellbeing"), then keep only sessions that had a "pre" game measure of mood at time zero and at least one more mood report.

```{r}

dat <- dat |> filter(prompt == "Wellbeing") |> select(-prompt)

# Count of data with all wellbeing observations
dat_total_wellbeing <- dat |> 
  summarise(
    n_total = n(),
    n_players = length(unique(pid))
  )

dat <- dat |> 
  mutate(
    session_has_pre = duration[1] == 0,
    session_has_post = sum(!is.na(mood)) > 1,
    .by = c(pid, session)
  )

dat |> 
  count(session_has_pre, session_has_post) |> 
  mutate(p = n / sum(n))

dat <- dat |> 
  filter(session_has_pre, session_has_post) |> 
  select(!contains("_has_"))
```

```{r}
bind_rows(
  "Total" = dat_total_wellbeing,
  "This study" = dat |> 
    summarise(
      n_total = n(),
      n_players = length(unique(pid))
    ),
  .id = "Data"
)
```

The data look like in @tbl-data.

```{r}
#| label: tbl-data
#| tbl-cap: First few rows of the data.

dat |> 
  slice(1:10) |> 
  gt() |> 
  fmt_datetime(time) |> 
  cols_align("left") |> 
  opt_stylize(6)
```

We then summarise the raw values of session durations and numbers of sessions in @fig-sessions.

```{r}
#| label: fig-sessions
#| fig-height: 5
#| fig-cap: '**A**. Raw session durations. **B**. Session durations after trimming top 1%. **C**. Raw numbers of sessions per participant. **D**. As C, but after trimming participants in the top 1% on number of sessions.'

p0 <- dat |> 
  ggplot() +
  scale_x_continuous(
    expand = expansion(c(0.01, 0.1)),
  ) +
  scale_y_continuous(
    expand = expansion(c(0.01, 0.1)),
  ) +
  geom_histogram(
    col = "dodgerblue4",
    fill = "dodgerblue1",
    linewidth = .25
  )

p_durations <- p0 +
  aes(duration) +
  labs(x = "Session duration (minutes)", y = "Observations")

p_durations_99 <- p_durations %+%
  filter(dat, duration < quantile(duration, .99))

dat_sessions <- dat |> 
  summarise(
    sessions = max(session),
    .by = pid
  )

p_sessions <- p0 %+% 
  dat_sessions +
  aes(sessions) +
  labs(x = "Sessions", y = "Participants")

p_sessions_99 <- p_sessions %+% 
  filter(p_sessions$data, sessions < quantile(sessions, .99))

(p_durations | p_durations_99) / 
  (p_sessions | p_sessions_99) +
  plot_annotation(tag_level = "A")
```

Based on these very long tails of implausible values (especially session duration), we trim at 99%.

```{r}
dat <- dat |> 
  filter(duration < quantile(duration, .99)) |> 
  right_join(
    dat_sessions |> 
      filter(sessions < quantile(sessions, .99)) |> 
      select(pid)
  )
```

Here is a scatterplot of the raw trimmed data.

```{r}
#| label: fig-scatter
#| fig-cap: Scatterplot of mood on duration

dat |> 
  ggplot() +
  scale_y_continuous(
    "Mood",
    expand = expansion(0.01)
  ) +
  scale_x_continuous(
    "Session duration (minutes)",
    expand = expansion(0.01)
  ) +
  aes(duration, mood) +
  geom_point(
    col = "dodgerblue",
    alpha = .2,
    size = .8
  )
```

We then save this dataset

```{r}
dat |> 
  write_rds("data/mood.rds")
```

# Descriptives

## Demographics

```{r}
#| echo: true
#| include: true

read_csv(
  "data/demographics.csv", 
  col_select = c(pid, country, gender, age)
) |> 
  filter(pid %in% unique(dat$pid)) |> 
  write_rds("data/demographics.rds")
demographics <- read_rds("data/demographics.rds")
demographics
```

```{r}
#| label: tbl-demographics
#| tbl-cap: Demographics
#| tbl-subcap: 
#|   - "Countries of origin"
#|   - "Gender responses"
#| layout-ncol: 2

count(demographics, country, sort = TRUE) |> 
  mutate(p = percent(n / sum(n), .1)) |> 
  gt()
demographics |> 
  mutate(gender = fct_lump_min(gender, 2)) |> 
  count(gender, sort = TRUE) |> 
  mutate(p = percent(n / sum(n), .1)) |> 
  gt()
```

```{r}
#| label: fig-age
#| fig-height: 3
#| fig-cap: Age distribution

p0 %+%
  demographics +
  aes(age)
```

# Modelling strategy

Here's our thinking of the modelling strategy at this point. Although this is not a pre-post experiment, it helps to label the setup as such:

- We rescale the original mood values from 0-1000 to 0-1 for easier interpretation
- We model mood reports for row *i*, person *j*, and session *k* with a normal distribution censored at 0 & 1
- We model the mean $\eta_{ijk}$ on an intercept (mood at pre-measure) and slope (mood difference btwn post and pre) with population level effects, and session and person deviations that are multivariate normal distributed
- We also model the residual deviation: random intercepts by person and probably by session too.

Here is a sketch of this analysis for one person

```{r}
#| label: fig-example
#| fig-cap: Sketch of the modelling strategy for one participant
#| fig-height: 3
library(lme4)
library(latex2exp)

dat <- dat |> 
  mutate(
    i = row_number(),
    i = factor(i == 1, levels = c(TRUE, FALSE), labels = c("pre", "post")),
    .by = c(pid, session)
  )

tmp <- dat |> 
  filter(pid == "p10004") |> 
  mutate(session = str_glue("{pid} (session: {session})"))
x <- lmer(
  mood ~ 0 + i + (0 + i | session), data = tmp 
)

# Ugly code >>> no code
newx <- tmp |> 
  data_grid(duration = seq_range(duration, 10), i, session)
newx$.fitted <- predict(x, newdata = newx)

newy <- tmp |> 
  data_grid(duration = seq_range(duration, 10), i)
newy$.fitted <- predict(x, newdata = newy, re.form = NA)
newy <- newy |> 
  mutate(session = "p10004 (all data)")
tmp2 <- tmp |> 
  mutate(session = "p10004 (all data)")

tmp <- bind_rows(tmp, tmp2)
newx <- bind_rows(newx, newy)

labels <- coef(x)$session |> 
  as.data.frame() |> 
  rownames_to_column("session") |> 
  mutate(duration = 50)

labels2 <- fixef(x) |> t() |> 
  data.frame() |> 
  mutate(session = "p10004 (all data)", duration = 50)

labels <- bind_rows(labels, labels2)

tmp |> 
  mutate(session = fct_inorder(session)) |> 
  ggplot() +
  aes(duration, mood, col = i) +
  scale_color_brewer(
    "Pre-session measure",
    palette = "Set1",
    aesthetics = c("color", "fill")
  ) +
  scale_x_continuous(
    "Session duration (mins)"
  ) +
  scale_y_continuous(
    "Mood",
    limits = c(0, 1)
  ) +
  geom_line(
    data = newx |> mutate(session = fct_inorder(session)),
    aes(y = .fitted)
  ) +
  geom_segment(
    data = labels |> mutate(session = fct_inorder(session)),
    aes(x = duration, xend = duration, y = ipre, yend = ipost, col = "diff"),
    arrow = arrow(length = unit(5, "pt"), ends = "both", type = "closed")
  ) +
  geom_point() +
  labs(caption = "mood ~ 1 + post + (1 + post | session)") +
  facet_wrap("session", nrow = 1) +
  theme(legend.position = "bottom")
```
